```{r}
library(tidyverse)
library(cowplot)
```

Quick exploration of possible mechanisms to account for differences in distributions seem with myRF data.

Assume that the measured diameters are the result of multiplicative interactions between two processes, each of which generates normally distributed outputs.

```{r}
mean1 <- 1
mean2 <- 1
sd1_con <- 1
sd1_mut <- 0.5
sd2 <- 1

df <- tibble(proc1_con = rnorm(100, mean1, sd1_con),
             proc1_mut = rnorm(100, mean1, sd1_mut),
             proc2 = rnorm(100, mean2, sd2),
             out_con = proc1_con * proc2,
             out_mut = proc1_mut * proc2) %>%
    gather("group", "values", 1:5)


(proc1_gg <- ggplot(df, aes(factor(group, c('proc1_con','proc1_mut', 'proc2', 'out_con', 'out_mut')), values)) +
     geom_violin())

ggsave('Plots/simulations.jpeg', proc1_gg)
```

In the control group, the distribution of axon diameters (out_con) is a result of multiplying two input processes (proc1_con x proc2).

In the mutant group, the distribution of axon diameters (out_mut) is also a result of multiplying two input processes (proc1_mut x proc2), but now the variability in the first input process is halved. 

